---
title: "Data Visualizations"
output: html_notebook
---

## Class catchup

```{r}
library(tidyverse)
library(rlang)
library(dbplyr)
library(dbplot)
# Class catchup
con <- DBI::dbConnect(odbc::odbc(), "Postgres Dev")
airports <- tbl(con, in_schema("datawarehouse", "airport")) 
flights <- tbl(con, in_schema("datawarehouse", "vflight"))
carriers <- tbl(con, in_schema("datawarehouse", "carrier"))
```

## 4.1 - Simple plot

1. Use `collect()` bring back the aggregated results into a "pass-through" variable called `by_month`
```{r, operations}
by_month <- flights %>%
  group_by(month) %>%
  tally() %>%
  collect() 

head(by_month)
```

2. Plot results using `ggplot2`
https://github.com/tidyverse/ggplot2/issues/2377

```{r}
library(ggplot2)

ggplot(by_month) +
  geom_line(aes(x = month, y = n)) 
```

## 4.2 - Plot in one code segment

1. Using the code from the previous section, create a single piped code set which also creates the plot
```{r}

```

2. Change the aggregation to the average of `arrdelay`.  Tip: Use `x` as the summarize variable 
```{r}

```

3. Plot the average distance. Copy the code from the previous exercise and change the variable
```{r}

```

## 4.3 - Plot specific data segments

1. Start with getting the top 5 carriers
```{r}
flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(5) 
```

2. Pipe the top 5 carriers to a plot. Use `geom_col`
```{r}
flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  mutate(n = as.numeric(n)) %>%
  arrange(desc(n)) %>%
  head(5) %>%
  collect() %>%
```

3. Improve the plot's look
```{r}
flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  mutate(n = as.numeric(n)) %>%
  arrange(desc(n)) %>%
  head(5) %>%
  collect() %>%
```

## 4.4 - Two or more queries

1. Use `pull()` to get the top 5 carriers loaded in a vector
```{r}
top5 <- flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(5) %>%
  pull(uniquecarrier)

top5
```

2. Use `%in%` to pass the `top5` vector to a filter
```{r}
flights %>%
  filter(uniquecarrier %in% top5) 
```

3. Group by *uniquecarrier* and get the average arrival delay
```{r}

```

4. Copy the final `ggplot()` code from the *Plot specific segment* section. Update the `y` labs.
```{r}

```

## 4.5 - Visualize using `dbplot`

1. Install and load `dbplot`
```{r, eval = FALSE}
library(dbplot)
```

2. Create a line plot using the helper function `dbplot_line()`

```{r}
flights %>%
  dbplot_line(month)
```

3. Update the plot's labels
```{r}
flights %>%
  dbplot_line(month) +

```

## 4.6 - Plot a different aggregation

1. Plot the average departure delay by day of week
```{r}
flights %>%
  dbplot_bar(dayofweek, )
```

2. Change the day numbers to day name labels
```{r}
flights %>%
  dbplot_bar(dayofweek, ) +
```

## 4.7 - Create a histogram

1. Use the `dbplot_histogram()` to build the histogram
```{r}
flights %>%
  dbplot_histogram(distance)
```

2. Adjust the `binwidth` to 300

```{r}
flights %>%
  dbplot_histogram(distance, )
```

## 4.8 - Raster plot

1. Use a `dbplot_raster()` to visualize `deptime` versus `arrtime`
```{r}
flights %>%
  
```

2. Change the plot's resolution to 500
```{r}
  
```


## 4.9 - Using the `calculate` functions

1. Use the `db_compute_raster()` function to get the underlying results that feed the plot
```{r}
departure <- flights %>%
  db_compute_raster(deptime, arrtime)

departure
```

2. Plot the results "manually"
```{r}
departure %>%
  filter(`n()` > 1000) %>%
  ggplot() +
  geom_raster(aes(x = deptime, y = arrtime, fill = `n()`))
```

## 4.10 - Under the hood 

1. Use the `db_bin()` command to see the resulting tidy eval formula
```{r}
db_bin(field)
```

2. Use `translate_sql()` and `simulate_odbc_postgresql()` to see an example of what the resulting SQL statement looks like
```{r}

```

3. Disconnect from the database

```{r}
dbDisconnect(con)
```
