---
Title: "Modeling"
output: html_notebook
---

## Class catchup

```{r}
library(tidyverse)
library(DBI)
library(dbplyr)
library(dbplot)
library(tidypredict)
con <- DBI::dbConnect(odbc::odbc(), "Postgres Dev")
airports <- tbl(con, in_schema("datawarehouse", "airport")) 
table_flights <- tbl(con, in_schema("datawarehouse", "flight"))
carriers <- tbl(con, in_schema("datawarehouse", "carrier"))
set.seed(100)
```

## 5.1 - SQL Native sampling 

1. Use `build_sql()` and `remote_query()` to combine a the `dplyr` command with a custom SQL statement
```{r}
sql_sample <- dbGetQuery(con, build_sql(remote_query(table_flights), " TABLESAMPLE SYSTEM (0.1)"))
```

2. Preview the sample data
```{r}
sql_sample
```

3. Test the efficacy of the sampling with a plot
```{r}
dbplot_histogram(sql_sample, distance)
```

## 5.2 - Sample with ID

1. Use `max()` to get the upper limit for *flightid*
```{r}
limit <- table_flights %>%
  summarise(
    max = max(flightid, na.rm = TRUE),
    min = min(flightid, na.rm = TRUE)
  ) %>%
  collect()
```

2. Use `sample` to get 0.1% of IDs
```{r}
sampling <- sample(
  limit$min:limit$max, 
  round((limit$max -limit$min) * 0.001))
```

3. Use `%in%` to match the sample IDs in the *flight* table
```{r}
id_sample <- table_flights %>%
  filter(flightid %in% sampling) %>%
  collect()
```

 Verify sample with a histogram
```{r}
dbplot_histogram(id_sample, distance)
```

## 5.3 - Sample manually

1. Create a filtered dataset for with 1 month of data
```{r}
db_month <- table_flights %>%
  filter(month == 1)
```

2. Get the row count
```{r}
rows <- as.integer(pull(tally(db_month)))
```

3. Use `row_number()` to create a new column to number each row
```{r}
db_month <- db_month %>%
  mutate(row = row_number()) 
```

4. Create a random set of 600 numbers, limited by the number of rows
```{r}
sampling <- sample(1:rows, 600)
```

5. Use `%in%` to filter the matched sample row IDs with the random set
```{r}
db_month <- db_month %>%
  filter(row %in% sampling)
```

6. Verify number of rows
```{r}
tally(db_month)
```

7. Create a function with the previous steps, but replacing the month number with an argument.  Collect the data at the end
```{r}
sample_segment <- function(x, size = 600) {
  db_month <- table_flights %>%
    filter(month == x)
  rows <- as.integer(pull(tally(db_month)))
  db_month <- db_month %>%
    mutate(row = row_number())
  sampling <- sample(1:rows, size)
  db_month %>%
    filter(row %in% sampling) %>%
    collect()
}
```

8. Test the function
```{r}
head(sample_segment(3), 100)
```

9. Use `map_df()` to run the function for each month
```{r}
strat_sample <- 1:12 %>%
  map_df(~sample_segment(.x))
```

10. Verify sample with a histogram
```{r}
dbplot_histogram(strat_sample, distance)
```

## 5.4 - Create a model & test

1. Prepare a model data set
```{r}
model_data <- strat_sample %>%
    mutate(
    season = case_when(
      month >= 3 & month <= 5  ~ "Spring",
      month >= 6 & month <= 8  ~ "Summmer",
      month >= 9 & month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter"
    )
  ) %>%
  select(arrdelay, season, depdelay) 
  
```

2. Create a simple `lm()` model
```{r}
model_lm <- lm(arrdelay ~ . , data = model_data)
summary(model_lm)
```

3. Create a test data set by combining the sampling and model data set routines. Set the `sample_segment()` `size` to 100
```{r}

```

4. Run a simple routine to check accuracy 
```{r}
test_sample %>%
  mutate(p = predict(model_lm, test_sample),
         over = abs(p - arrdelay) < 10) %>%
  group_by(over) %>% 
  tally() %>%
  mutate(percent = round(n / sum(n), 2))
```

## 5.5 - Score inside database

1. Load the library, and see the results of passing the model as an argument to `tidypredict_fit()` 
```{r}
library(tidypredict)

tidypredict_fit(model_lm)
```

2. Use `tidypredict_sql()` to see the resulting SQL statement
```{r}
tidypredict_sql(model_lm, con)
```

3. Run the prediction inside `dplyr`
```{r}
table_flights %>%
  filter(month == 2,
         dayofmonth == 1) %>%
    mutate(
    season = case_when(
      month >= 3 & month <= 5  ~ "Spring",
      month >= 6 & month <= 8  ~ "Summmer",
      month >= 9 & month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter"
    )
  ) %>%
  select( season, depdelay) %>%
  tidypredict_to_column(model_lm) %>%
  head()
```

4. View the SQL behind the `dplyr` command. Use `remote_query()`
```{r}

```

5. Compare predictions to ensure results are within range
```{r}
test <- tidypredict_test(model_lm)
test
```

6. View any records that exceeded the threshold
```{r}
test$raw_results %>%
  filter(fit_threshold)
```

## 5.6 - Parsed model

1. Use the `parse_model()` function to see how `tidypredict` interprets the model
```{r}
pm <- parse_model(model_lm)
pm
```

2. Verify that the resulting table can be used to get the fit formula
```{r}
tidypredict_test(model_lm)
```

3. Using `write_csv()`, save the parsed model for later use
```{r}

``` 

4. Disconnect from the database

```{r}
dbDisconnect(con)
```


