---
title: "Intro to dashboards"
output: html_notebook
---

## Class catchup

```{r}
library(tidyverse)
library(dbplyr)
library(DBI)
library(dbplot)

# Class catchup
con <- DBI::dbConnect(odbc::odbc(), "Postgres Dev")
airports <- tbl(con, in_schema("datawarehouse", "airport"))
flights <- tbl(con, in_schema("datawarehouse", "flight"))
carriers <- tbl(con, in_schema("datawarehouse", "carrier"))
```

## 9.1 - Basic structure

1. Create and preview a simple `shinydashboard` 
```{r, eval = FALSE}
library(shinydashboard)
library(shiny)
library(DT)

ui <- dashboardPage(
  dashboardHeader(title = "Quick Example"),
  dashboardSidebar(selectInput("select", "Selection", c("one", "two"))),
  dashboardBody(
    valueBoxOutput("total"),
    dataTableOutput("monthly")
  )
)

server <- function(input, output, session) {
  output$total <- renderValueBox(valueBox(100, subtitle = "Flights"))
  output$monthly <- renderDataTable(datatable(mtcars))
}

shinyApp(ui, server)
```

## 9.2 - Dropdown data


1. Use `purrr` to create a list with the correct structure for the `shiny` drop down
```{r}
airline_list <- carriers %>%  
  select(carrier, carriername) %>%   # In case more fields are added
  collect()  %>%                     # All would be collected anyway
  split(.$carriername) %>%           # Create a list item for each name
  map(~.$carrier)                    # Add the carrier code to each item

head(airline_list)
```

2. In the app code, replace `c("one", "two", "three")` with `airline_list`

```{r, eval = FALSE}
# Goes from this:
dashboardSidebar(selectInput("select", "Selection", c("one", "two"))),
# To this:
dashboardSidebar(selectInput("select", "Selection", airline_list)),
```

3. Re-run the app

## 9-3 - Update dashboard items

1. Save the base "query" to a variable. It will contain a carrier selection. To transition into `shiny` programming easier, the variable will be a function.
```{r}
base_dashboard <- function(){
  flights %>%
    filter(uniquecarrier == "DL")
  }

head(base_dashboard())
```

3. Use the base query to figure the number of flights for that carrier
```{r}
base_dashboard() %>%
  tally() %>% 
  pull()
```

4. In the app, remove the `100` number and pipe the `dplyr` code into the valueBox() function
```{r, eval = FALSE}
# Goes from this:
  output$total <- renderValueBox(valueBox(100, subtitle = "Flights"))
# To this:
  output$total <- renderValueBox(
    base_dashboard() %>%
      tally() %>% 
      pull() %>%
      valueBox(subtitle = "Flights"))
```

5. Create a table with the month name and the number of flights for that month 
```{r}
base_dashboard() %>%
  group_by(month) %>%
  tally() %>%
  collect() %>%
  mutate(n = as.numeric(n)) %>%
  rename(flights = n) %>%
  arrange(month)

```

6. In the app, replace `head(mtcars)` with the piped code, and re-run the app
```{r, eval = FALSE}
# Goes from this:
  output$monthly <- renderTable(head(mtcars))
# To this:
  output$monthly <- renderDataTable(datatable(
    base_dashboard() %>%
      group_by(month) %>%
      tally() %>%
      collect() %>%
      mutate(n = as.numeric(n)) %>%
      rename(flights = n) %>%
      arrange(month)))
```

## 9.4 - Integrate the dropdown

1. In the original `base_dashboard()` code, replace `function` with `reactive`, and `"DL"` with `input$select`
```{r, eval = FALSE}
# Goes from this
base_dashboard <- function(){
flights %>%
  filter(uniquecarrier == "DL")}
# To this
base_dashboard <- reactive({
  flights %>% 
    filter(uniquecarrier == input$select)})
```

2. Insert the new code right after the `server <- function(input, output, session)` line. The full code should now look like this:
```{r, eval = FALSE}

ui <- dashboardPage(
  dashboardHeader(title = "Quick Example"),
  dashboardSidebar(selectInput("select", "Selection", airline_list)),
  dashboardBody(
    valueBoxOutput("total"),
    dataTableOutput("monthly")
  )
)

server <- function(input, output, session) {
  base_dashboard <- reactive({
    flights %>%
      filter(uniquecarrier == input$select)
  })
  output$total <- renderValueBox(
    base_dashboard() %>%
      tally() %>%
      pull() %>%
      valueBox(subtitle = "Flights")
  )
  output$monthly <- renderDataTable(datatable(
    base_dashboard() %>%
      group_by(month) %>%
      tally() %>%
      collect() %>%
      mutate(n = as.numeric(n)) %>%
      rename(flights = n) %>%
      arrange(month)
  ))
}
shinyApp(ui, server)
```

9. Re-run the app

10. Disconnect form database
```{r}
dbDisconnect(con)
```

