
```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
```


# `dplyr` Basics

```{r, include = FALSE}

library(dplyr)
library(dbplyr)
library(DBI)
```

## Create a table variable

*Basics to how to point a variable in R to a table or view inside the database*


1. Load the `dplyr` and `dbplyr` libraries
```{r, dplyr}
library(dplyr)
library(dbplyr)
```

2. *(Optional)* Open a connection to the database if it's currently closed
```{r}
library(DBI)

con <- dbConnect(odbc::odbc(), "Postgres Dev")
```


3. Use the `tbl()` and `in_schema()` functions to create a reference to a table
```{r}
tbl(con, in_schema("datawarehouse", "airport"))
```

4. Load the reference, not the table data, into a variable
```{r}
airports <- tbl(con, in_schema("datawarehouse", "airport"))
```


5. Call the variable to see preview the data in the table
```{r}
airports
```

6. Set up the pointers to the other of the tables
```{r}
flights <- tbl(con, in_schema("datawarehouse", "flight"))
carriers <- tbl(con, in_schema("datawarehouse", "carrier"))
```

## Basic aggregation
*A couple of `dplyr` commands that run in-database*

1. How many records are in the **airport** table?
```{r}
tbl(con, in_schema("datawarehouse", "vflight"))  %>%
  group_by(month) %>%
  tally()
```

2. What is the average character length of the airport codes? How many characters is the longest and the shortest airport name?
```{r}
airports %>%
  summarise(
    avg_airport_length = mean(length(airport), na.rm = TRUE),
    max_airport_name = max(length(airportname), na.rm = TRUE),
    min_airport_name = min(length(airportname), na.rm = TRUE),
    total_records = n()
  )
```

**Additional exercises:**

1. How many records are in the **carrier** table?

2. How many characters is the longest **carriername**?

## Under the hood 
* Use `show_query()` to preview the SQL statement that will be sent to the database*

1. SQL statement that actually runs when we ran `airports` as a command
```{r}
show_query(airports)
```

2. Easily view the resulting query by adding `show_query()` in another piped command
```{r}
carriers %>%
  summarise(n())
```

3. Run the same for last exercise in the previous section
```{r}
airports %>%
  summarise(
    avg_airport_length = mean(length(airport), na.rm = TRUE),
    max_airport_name = max(length(airportname), na.rm = TRUE),
    min_airport_name = min(length(airportname), na.rm = TRUE),
    total_records = n()
  ) %>%
  show_query()
```

## Un-translated R commands
*Review of how `dbplyr` handles R commands that have not been translated into a like-SQL command*

1. Preview how `Sys.time()` is translated
```{r}
airports %>%
  mutate(today = Sys.time()) %>%
  show_query()
```

2. Use PostgreSQL's native commands, in this case `now()`
```{r}
airports %>%
  mutate(today = now()) %>%
  show_query()
```

3. Run the `dplyr` code to confirm it works
```{r}
airports %>%
  mutate(today = now()) %>%
  select(today) %>%
  head()
```


```{r, include = FALSE}
dbDisconnect(con)
```
