```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
```


# Data transformation

```{r, include = FALSE}

library(dplyr)
library(dbplyr)

# Class catchup
con <- DBI::dbConnect(odbc::odbc(), "Postgres Dev")
airports <- tbl(con, in_schema("datawarehouse", "airport"))
flights <- tbl(con, in_schema("datawarehouse", "flight"))
carriers <- tbl(con, in_schema("datawarehouse", "carrier"))

```

## Group and sort records

*Learn how to use `group_by()` and `arrange()` to better understand aggregated data*


1. How many flights are there per month?
```{r, transformation}
flights %>%
  group_by(month) %>%
  tally() 
```

2. Order the results by the month number
```{r}
flights %>%
  group_by(month) %>%
  tally() %>%
  arrange(month)
```

3. Order the results by the number of flights, starting with the month with most flights by using `desc()` inside the `arrange()` command
```{r}
flights %>%
  group_by(month) %>%
  tally() %>%
  arrange(desc(n)) 
```

## Answering questions with `dplyr`
*Quick review of how to translate questions into `dplyr` code*

1. Which are the top 4 months with the most flight activity?
```{r}
flights %>%
  group_by(month) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(4)
```

2. What were the top 5 calendar days with most flight activity?
```{r}
flights %>%
  group_by(month, dayofmonth) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(5)
```

3. Which are the top 5 carriers (airlines) with the most flights?

```{r}
flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(5)
```

4. Figure the percent ratio of flights per month
```{r}
flights %>%
  group_by(month) %>%
  tally() %>%
  arrange(desc(n)) %>%
  mutate(percent = n/sum(n, na.rm = TRUE))
```

## Aggregate mulitple columns
*Practice using `summarise _` functions*


1. Use `summarise_all()` to send the same function to all fields
```{r}
flights %>%
  select(depdelay, arrdelay) %>%
  summarise_all(mean, na.rm = TRUE)
```

2. Use `summarise_at()` to pre-select the fields that will receive the function
```{r}
flights %>%
  summarise_at(c("depdelay", "arrdelay"),mean, na.rm = TRUE)
```

3. Use `summarise_if()` to summarize only if the field meets a criterion
```{r}
flights %>%
  summarise_if(is.numeric,mean, na.rm = TRUE)
```

4. Combine with `group_by()` to create more complex results
```{r}
flights %>%
  select(month, depdelay, arrdelay) %>%
  group_by(month) %>%
  summarise_all(mean, na.rm = TRUE)
```

## View record level data
*Important tips to record preview data*


How many flights in July 18th were one or more hours late?
```{r}
flights %>%
  filter(depdelay >= 60,
         month == 7,
         dayofmonth == 18) %>%
  tally() 
```


1. Use `filter()` to retrieve only the needed data, and `head()` to limit the preview even further.
```{r}
flights %>%
  filter(depdelay >= 60,
         month == 7,
         dayofmonth == 18) %>%
  head(100)
```

2. Use `collect()` and `View()` to preview the data in the IDE. Make sure to **always** limit the number of returned rows.
```{r, eval = FALSE}
flights %>%
  filter(depdelay >= 60,
         month == 7,
         dayofmonth == 18) %>%
  collect() %>%
  head(100) %>%
  View("my_preview")
```

## Case statements
*See how to use the flexibility of case statements for special cases*

1. Use `case_when()` to bucket each month one of four seasons
```{r}
flights %>%
  mutate(
    season = case_when(
      month >= 3 && month <= 5  ~ "Spring",
      month >= 6 && month <= 8  ~ "Summmer",
      month >= 9 && month <= 11 ~ "Fall",
      TRUE ~ "Winter"
    )
  ) %>%
  group_by(season) %>%
  tally()
```

2. Add a specific case for "Winter"
```{r}
flights %>%
  mutate(
    season = case_when(
      month >= 3 && month <= 5  ~ "Spring",
      month >= 6 && month <= 8  ~ "Summmer",
      month >= 9 && month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter"
    )
  ) %>%
  group_by(season) %>%
  tally()
```

3. Append an entry for Monday at the end of the case statement
```{r}
flights %>%
  mutate(
    season = case_when(
      month >= 3 && month <= 5  ~ "Spring",
      month >= 6 && month <= 8  ~ "Summmer",
      month >= 9 && month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter",
      dayofweek == 1 ~ "Monday"
    )
  ) %>%
  group_by(season) %>%
  tally()
```

4. Move the "Monday" entry to the top of the case statement
```{r}
flights %>%
  mutate(
    season = case_when(
      dayofweek == 1 ~ "Monday",
      month >= 3 && month <= 5  ~ "Spring",
      month >= 6 && month <= 8  ~ "Summmer",
      month >= 9 && month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter"
    )
  ) %>%
  group_by(season) %>%
  tally()
```


```{r, include = FALSE}
dbDisconnect(con)
```
