# Data Visualizations


```{r, include = FALSE}
knitr::opts_chunk$set(cache = FALSE)

library(dplyr)
library(dbplyr)
library(dbplot)
library(ggplot2)
# Class catchup
con <- DBI::dbConnect(odbc::odbc(), "Postgres Dev")
airports <- tbl(con, in_schema("datawarehouse", "airport")) %>%
  mutate(crsarrtime = crsarrtime)


flights <- tbl(con, in_schema("datawarehouse", "flight"))
carriers <- tbl(con, in_schema("datawarehouse", "carrier"))



```



## Simple plot
*Practice pushing the calculations to the database*

1. Use `collect()` bring back the aggregated results into a "pass-through" variable called `by_month`
```{r}
by_month <- flights %>%
  group_by(month) %>%
  tally() %>%
  mutate(n = as.numeric(n)) %>%
  collect() 

head(by_month)
```

2. Plot results using `ggplot2`

```{r}
library(ggplot2)

ggplot(by_month) +
  geom_line(aes(x = month, y = n)) 
```

## Plot in one code segment
*Practice going from `dplyr` to `ggplot2` without using pass-through variable, great for EDA*

1. Using the code from the previous section, create a single piped code set which also creates the plot
```{r}
flights %>%
  group_by(month) %>%
  tally() %>%
  mutate(n = as.numeric(n)) %>%
  collect() %>%
  ggplot() +                        # < Don't forget to switch to `+`
  geom_line(aes(x = month, y = n)) 
```

2. Change the aggregation to the average of `arrdelay`.  Tip: Use `x` as the summarize variable 
```{r}
flights %>%
  group_by(month) %>%
  summarise(x = mean(arrdelay)) %>%
  mutate(x = as.numeric(x)) %>%
  collect() %>%
  ggplot() +
  geom_line(aes(x = month, y = x)) 
```

3. Plot the average distance 
```{r}
flights %>%
  group_by(month) %>%
  summarise(x = mean(distance)) %>%
  mutate(x = as.numeric(x)) %>%
  collect() %>%
  ggplot() +
  geom_line(aes(x = month, y = x)) 
```


## Plot specific data segments
*Combine skills from previous units to create more sophisticated plots*

1. Start with getting the top 5 carriers
```{r}
flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(5) 
```

1. Start with getting the top 5 carriers
```{r}
flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  mutate(n = as.numeric(n)) %>%
  arrange(desc(n)) %>%
  head(5) %>%
  collect() %>%
  ggplot() +
    geom_col(aes(x = uniquecarrier, y = n))
```

1. Improve the plot's look
```{r}
flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  mutate(n = as.numeric(n)) %>%
  arrange(desc(n)) %>%
  head(5) %>%
  collect() %>%
  ggplot() +      #Don't forget to switch to `+`
    geom_col(aes(x = uniquecarrier, y = n, fill = n)) +  #Add fill
    theme(legend.position="none") +    # Turn legend off
    coord_flip() +  # Rotate cols into rows
    labs(title = "Top 5 Carriers", 
         subtitle = "Source: Datawarehouse",
         x = "Carrier Name", 
         y = "# of Flights")
    
```

## 2 or more queries
*Learn how to use `pull()` to pass a set of values to be used on a secondary query*

1. Use `pull()` to get the top 5 carriers loaded in a vector

```{r}
top5 <- flights %>%
  group_by(uniquecarrier) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(5) %>%
  pull(uniquecarrier)

top5
```

2. Use `%in%` to pass the `top5` vector to a filter
```{r}
flights %>%
  filter(uniquecarrier %in% top5) 
```

3. Group by carrier and get the average arrival delay
```{r}
flights %>%
  filter(uniquecarrier %in% top5) %>%
  group_by(uniquecarrier) %>%
  summarise(n = mean(arrdelay))
```

4. Copy the final `ggplot()` code from the *Plot specific segment* section. Update the `y` labs.
```{r}
flights %>%
  filter(uniquecarrier %in% top5) %>%
  group_by(uniquecarrier) %>%
  summarise(n = mean(arrdelay)) %>%
  # From previous section ----------------------------------------------
  collect() %>%
  ggplot() +      #Don't forget to switch to `+`
    geom_col(aes(x = uniquecarrier, y = n, fill = n)) +  #Add fill
    theme(legend.position="none") +    # Turn legend off
    coord_flip() +  # Rotate cols into rows
    labs(title = "Top 5 Carriers", 
         subtitle = "Source: Datawarehouse",
         x = "Carrier Name", 
         y = "Average Delay") 
```

## Visualize using `dbplot`
*Review how to use `dbplot` to make it easier to plot with databases*

1. Install and load `dbplot`
```{r, eval = FALSE}
install.packages("dbplot")
library(dbplot)
```

2. Create a line plot using the helper function `dbplot_line()`

```{r}
flights %>%
  dbplot_line(month)
```

3. Update the plot's labels
```{r}
flights %>%
  dbplot_line(month) +
  labs(title = "Monthly flights",
       x = "Month",
       y = "Number of flights") 
```

## Plot a different aggregation
*`dbplot` allows for aggregate functions, other than record count, to be used for plotting*

1. Plot the average departure delay by day of week
```{r}
flights %>%
  dbplot_bar(dayofweek, mean(depdelay))
```

2. Change the day numbers to day name labels
```{r}
flights %>%
  dbplot_bar(dayofweek, mean(depdelay)) +
    scale_x_continuous(labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat","Sun"),
                     breaks = 1:7)
```



## Create a histogram
*Use the package's function to easily create a histogram*


1. Use the `dbplot_histogram()` to build the histogram
```{r}
flights %>%
  dbplot_histogram(distance)
```


2. Adjust the `binwidth` to 300

```{r}
flights %>%
  dbplot_histogram(distance, binwidth = 300)
```

## Raster plot

1. Use a `dbplot_raster()` to visualize `deptime` versus `depdelay`
```{r}
flights %>%
  dbplot_raster(deptime, arrtime)
```

2. Change the plot's resolution to 500
```{r}
flights %>%
  dbplot_raster(deptime, arrtime, resolution = 500)
```


## Using the `calculate` functions

1. Use the `db_comptue_raster()` function to get the underlying results that feed the plot
```{r}
departure <- flights %>%
  db_compute_raster(deptime, arrtime)

departure
```

2. Plot the results "manually"
```{r}
departure %>%
  filter(`n()` > 1000) %>%
  ggplot() +
  geom_raster(aes(x = deptime, y = arrtime, fill = `n()`))
```


## Under the hood (II)
*Review how `dbplot` pushes histogram and raster calculations to the database*

1. Use the `db_bin()` command to see the resulting tidy eval formula
```{r}
db_bin(field)
```

2. Use `trasnlate_sql()` and `simulate_odbc_postgresql()` to see an example of what the resulting SQL statement looks like
```{r}
translate_sql(!! db_bin(field), con = simulate_odbc_postgresql())
```




