# Modeling

```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, cache = TRUE)
```

```{r, include = FALSE}

library(dplyr)
library(dbplyr)
library(dbplot)
library(ggplot2)
library(purrr)
library(DBI)
# Class catchup
con <- DBI::dbConnect(odbc::odbc(), "Postgres Dev")
airports <- tbl(con, in_schema("datawarehouse", "airport")) %>%
  mutate(crsarrtime = crsarrtime)

flights <- tbl(con, in_schema("datawarehouse", "flight"))
carriers <- tbl(con, in_schema("datawarehouse", "carrier"))


set.seed(100)
```

## SQL Native sampling 
*Use PostgreSQL TABLESAMPLE clause*

1. Use `build_sql()` and `remote_query()` to combine a the `dplyr` command with a custom SQL statement
```{r}
sample_table<-  dbGetQuery(con, build_sql(remote_query(flights), " TABLESAMPLE SYSTEM (0.1)"))
```

2. Preview the sample data
```{r, eval = FALSE}
View(sample_table, 1000)
```

3. Test the efficacy of the sampling with a plot
```{r}
dbplot_histogram(sample_table, distance)
```

## Sample manually
*Use `row_number()`, `sample()` and `map_df()` to create a sample data set*


1. Create a filtered dataset for with 1 month of data
```{r}
db_month <- flights %>%
  filter(month == 1)
```

2. Get the row count
```{r}
rows <- as.integer(pull(tally(db_month)))
```

3. Use `row_number()` to create a new column to number each row
```{r}
db_month <- db_month %>%
  mutate(row = row_number()) 
```

4. Create a random set of 600 numbers, limited by the number of rows
```{r}
sampling <- sample(1:rows, 600)
```

5. Use `%in%` to filter the matched sample row IDs with the random set
```{r}
db_month <- db_month %>%
  filter(row %in% sampling)
```

6. Verify number of rows
```{r}
tally(db_month)
```

7. Create a function with the previous steps, but replacing the month number with an argument.  Collect the data at the end
```{r}
sample_segment <- function(x, size = 600) {
  db_month <- flights %>%
    filter(month == x)
  rows <- as.integer(pull(tally(db_month)))
  db_month <- db_month %>%
    mutate(row = row_number())
  sampling <- sample(1:rows, size)
  db_month %>%
    filter(row %in% sampling) %>%
    collect()
}
```

8. Test the function
```{r}
head(sample_segment(3), 100)
```

9. Use `map_df()` to run the function for each month
```{r}
new_sample <- 1:12 %>%
  map_df(~sample_segment(.x))
```

10. Verify sample with a histogram
```{r}
dbplot_histogram(new_sample, distance)
```

## Create a model & test

1. Prepare a model data set
```{r}
model_data <- new_sample %>%
    mutate(
    season = case_when(
      month >= 3 & month <= 5  ~ "Spring",
      month >= 6 & month <= 8  ~ "Summmer",
      month >= 9 & month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter"
    )
  ) %>%
  select(arrdelay, season, depdelay) 
  
```

2. Create a simple `lm()` model
```{r}
model_lm <- lm(arrdelay ~ . , data = model_data)
summary(model_lm)
```

3. Create a test data set by combining the sampling and model data set routines
```{r}
test_sample <- 1:12 %>%
  map_df(~sample_segment(.x, 100)) %>%
    mutate(
    season = case_when(
      month >= 3 & month <= 5  ~ "Spring",
      month >= 6 & month <= 8  ~ "Summmer",
      month >= 9 & month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter"
    )
  ) %>%
  select(arrdelay, season, depdelay) 
  
```

4. Run a simple routine to check accuracy 
```{r}
test_sample %>%
  mutate(p = predict(model_lm, test_sample),
         over = abs(p - arrdelay) < 10) %>%
  group_by(over) %>% 
  tally() %>%
  mutate(percent = round(n / sum(n), 2))
```

## Score inside database
*Learn about tidypredict to run predictions inside the database*

1. Load the library, and see the results of passing the model as an argument to `tidypredict_fit()` 
```{r}
library(tidypredict)

tidypredict_fit(model_lm)
```

2. Use `tidypredict_sql()` to see the resulting SQL statement
```{r}
tidypredict_sql(model_lm, con)
```

3. Run the prediction inside `dplyr`
```{r}
flights %>%
  filter(month == 2,
         dayofmonth == 1) %>%
    mutate(
    season = case_when(
      month >= 3 & month <= 5  ~ "Spring",
      month >= 6 & month <= 8  ~ "Summmer",
      month >= 9 & month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter"
    )
  ) %>%
  select( season, depdelay) %>%
  tidypredict_to_column(model_lm) %>%
  head()
```

4. View the SQL behind the `dplyr` command
```{r}
flights %>%
  filter(month == 2,
         dayofmonth == 1) %>%
    mutate(
    season = case_when(
      month >= 3 & month <= 5  ~ "Spring",
      month >= 6 & month <= 8  ~ "Summmer",
      month >= 9 & month <= 11 ~ "Fall",
      month == 12 | month <= 2  ~ "Winter"
    )
  ) %>%
  select( season, depdelay) %>%
  tidypredict_to_column(model_lm) %>%
  remote_query()
```

6. Compare predictions to ensure results are within range
```{r}
test <- tidypredict_test(model_lm)
test
```

7. View the records that exceeded the threshold
```{r}
test$raw_results %>%
  filter(fit_threshold)
```

```{r, include = FALSE}
dbDisconnect(con)
```


